Compositional equivalences based on Open pNets

### 2020-04-17 2016-Weak-FH-Bisim/WeakBisim.tex
changes over skype

For this property to be valid even if the substitution does not operate on all variables,
we define the composition operation as follows: 
\TODO{version d eric reprise et corrigee}
%$(x_k\gets e_k)^{k\in K}\shortotimes \Post =  (x_k\gets e_k\subst{\Post})^{k\in K}$
%\ERIC{J'ai un probleme ici... quand $Post'$ agit sur des vars de $t$ qui ne sont pas substitutees par $Post$, alors les 2 defs de $\shortotimes$ ci-dessus ne sont pas equivalentes:}\\
%$F(x,z)(\subst{x\gets z}\shortotimes\subst{z\gets 1})=F(x,z)(\subst{x\gets 1})=F(1,z)$\\
%\ERIC{Alors que}\\
%$(F(x,z)\subst{x\gets z})\subst{z\gets 1})=F(z,z)\subst{z\gets 1}=F(1,1)$
%
%\ERIC{Une definition plus correcte serait peut-etre:}\\

$(x_{k}\gets e_{k})^{k\in K}\shortotimes (x'_{k'}\gets e'_{k'})^{k'\in K'} =  
(x_{k}\gets e_{k}\subst{(x'_{k'}\gets e'_{k'})^{k'\in K'}})^{k\in K} \cup (x'_{k'}\gets e'_{k'})^{k'\in K''}$\\
where $K''=\{k'\in K'|x'_{k'}\not\in\{x_k\}^{k\in K}\}$
%
%\LUDO{En fait comme on supose qu il y a pas de variable globale et donc que les post doivent affecter toutes les variables ce contre-exemple n est pas applicable! apres on peut definir comme ca si on veut ceinture et bretelles mais ca alourdit un peu la definitions}
%
%\ERIC{Heu Non... 
%  En pratique, dans la composition de posts de la regle WT3, les posts assignent bien toutes les variables de leur OT (strong), mais dans la composition, on fini par avoir des trous... Voir page 21-22. On est tombes tous les deux separemment, Rabea et moi, sur ce genre de cas.}
% 
%\LUDO{euuh ca doit etre lie qu fait qu il manque des ec leftarrowec dans les selfloop de l'OA de spec. Plutot que changer la def et les preuves je propose de rajouter la condition $vars(s')=dom(Post)$ pour tous les OAs ce qui ne change rien a part pour les selfloop que l on s autorisait apparemment a ecrire en omettant $x:=x$ ... au pire l'identite est un post toujours correct et valide}

@@

\begin{definition}[pLTS]
\label{pLTS}
A pLTS is a tuple
$\pLTS\triangleq\mylangle S,s_0, \to\myrangle$ where:
\begin{itemize}
\item[$\bullet$]
$S$ is a set of states.
\item[$\bullet$]
$s_0 \in S$ is the initial state.
%\item[$\bullet$]
 %Variables in
%$\iv(\alpha)$ are assigned by the action, other variables can be assigned
%by the additional assignments.
\item[$\bullet$] $\to \subseteq S \times L \times S$ is the transition relation and 
$L$ is the set of labels of the form
$\langle \alpha,~e_b,~(x_j\!:= {e}_j)^{j\in J}\rangle$,
where $\alpha \in\AlgA$ is a parameterised action, $e_b \in
\AlgB$ is a guard, and the variables $x_j$ 
are assigned the expressions $e_j\in \AlgE$.
If 
$s \xrightarrow{\langle \alpha,~e_b,~(x_j\!:= {e}_j)^{j\in
		J}\rangle} s'\in \to $ then 
% REMOVED BECAUSE USELESS: $\iv(\alpha)\!\subseteq\! \vars(s')$, 
		$\vars(\alpha)\backslash \iv(\alpha)\!\subseteq\! \vars(s)$, 
		$\vars(e_b)\!\subseteq\! \vars(s)\cup\vars(\alpha)$, and
		$\forall j\!\in\! J .\,\left(\vars(e_j)\!\subseteq\! \vars(s)\cup\vars(\alpha)\land 
		x_j\!\in\!\vars(s') \right)$. %,  and $s= s'\Rightarrow J=\emptyset$.
\hl{Additionally we require that:} $\{x_j|j\in J\} = \vars(s')$.
%\LUDO{I suggest to add: $\{x_j|j\in J\} = \vars(s')$ or $\{x_j|j\in J\} \cup \iv(\alpha)= \vars(s')$} 
\end{itemize}
\end{definition}

@@

\item[$\bullet$] $\symb{SV}_k^{k\in K}$ is a set of
  synchronisation vectors. $\forall k\!\in\! K,
  \symb{SV}_k\!=\!\SV{\alpha_{l}^{l\in I_k \uplus J_k}}{\alpha'_k}{e_k}$ \hl{where}
  $\alpha'_k\in \mathcal{A}_\P$, $I_k\subseteq I$, $J_k\subseteq J$,
  $\forall i\!\in\!
  I_k.\,\alpha_{i}\!\in\!\Sort(\pNet_i)$,  $\forall j\!\in\!
  J_k.\,\alpha_{j}\!\in\!\Sort_j$, and $\vars(\alpha'_k)\subseteq \bigcup_{l\in I_k\uplus 
  J_k}{\vars({\alpha_l})}$. The global action of a vector $\symb{SV}_k$ is
$\alpha'_k$. $e_k \in \AlgB$ is a guard associated to the vector such that
$\vars(e_k)\subseteq \bigcup_{l\in I_k\uplus J_k}{\vars({\alpha_l})}$.
\end{itemize}
Synchronisation vectors are identified modulo renaming of variables that appear in their 
action terms.
\end{definition}

@@

  \begin{itemize}
  \item The sort of a pNet is its signature, i.e. the set of actions it can
perform, where each action signature is an action 
label plus the arity of the action. In the definition of sorts, we do not need to 
distinguish
input variables and we remove the
\emph{input marker} (?) of variables.
\[
\begin{array}{l}
\Sortop(\mylangle S,s_0, \to\myrangle) = \{\alpha\subst{?x \gets x| 
x\in\symb{iv}(\alpha)}|s \xrightarrow{\langle \alpha,~e_b,~(x_j\!:= {e}_j)^{j\in
    J}\rangle} s'\in \to \} \\
\Sortop(\mylangle \set{\pNet}\!, %\pNet_i^{i\in I}, 
\set{\Sort},
\set{\symb{SV}}\myrangle)
=\{\alpha' |\, \SV{\set{\alpha}}{\alpha'}{e_b}\in\set{\symb{SV}}\}
\end{array}
\]

\item The set of variables of a pNet $P$, denoted $vars(P)$ is the set of variables of all states of all pLTSs that compose $P$.

\item
The set of holes $\Holes(\pNet)$ of a pNet is the indexes of the holes of the pNet 
itself plus the indexes of all the holes of its subnets.
It is defined inductively (we suppose those indexes 
disjoints):
  \[\begin{array}{l}
\Holes(\mylangle S,s_0, \to\myrangle) \!=\! \emptyset\\
\Holes(\mylangle \pNet_i^{i\in I}\!,\set{\Sort}, \set{\symb{SV}}\myrangle) 
=J\uplus{\displaystyle \bigcup_{i\in 
I}\Holes(\pNet_i)}\\
\forall i\in I.\, \Holes(\pNet_i)\cap J=\emptyset\\
\forall i_1,i_2\in I.\,i_1\neq i_2\Rightarrow  \Holes(\pNet_{i_1})\cap\Holes(\pNet_{i_2})=\emptyset
\end{array}\]
\item
The set of leaves of a pNet is the set of all pLTSs occurring in the structure, as an 
indexed family of the form $\Leaves(\pNet)= \mylangle \pNet_i \myrangle^{i \in L}$.
\[\begin{array}{l}
\Leaves(\mylangle S,s_0, \to\myrangle) \!=\!\emptyset\\%\{\mylangle S,s_0, \to\myrangle\}\\
\Leaves(\mylangle \pNet_i^{i\in I}\!,%Sort_j^{j\in J}
\set{\Sort}\!, \set{\symb{SV}}\myrangle) = {\displaystyle \biguplus_{i\in 
I}\Leaves(\pNet_i)\uplus\{i\mapsto \pNet_i|\pNet_i \text{ is a }\pLTS\}}
\end{array}\]
\end{itemize}

@@

\subsection{Open Automata}
 Open automata (OA) are not composition structures but they are made of transitions that are dependent of the actions of the holes, and they can reason on variables (potentially with only symbolic values). Formaly, each state of an OA has a set of \emph{state variables}, and all state variable sets are disjoint.
%\TODO{adopt a uniform notation for open transitions, almost each instance has a 
%different 
%notation! I suggest p,l,pr,po using \{\} for p,l,po as they are sets}
\begin{definition}[Open transitions]\label{def:OT}
	\label{def:OpenTransitions}
	An \emph{open transition} (OT) over a
	set $J$ of holes with sorts $\Sort_j^{j\in J}$ and a set of states $\mathcal{S}$ is 
	a structure of the form:	
	\begin{mathpar}
	\openrule
	{	\beta_j^{j\in J'}, \Pred, \Post}
	{s \OTarrow {\alpha}s'}
	\end{mathpar}
	Where $J'\subseteq J$, $s, s'\in\mathcal{S}$ and $\beta_j$
        is a transition of the hole $j$, with $\beta_j\in\Sort_j$. $\alpha$ is an action 
        label denoting the resulting action of this open transition.
        We call \emph{source variables} of an OT the union of all variables
        in the different terms
        $\beta_j$ and $\alpha$, and the state variables of $s$.
        \Pred\ is a predicate over the source variables. \Post\ is a set of 
	assignments that are effective \emph{after the open transition}, they are
        represented as a substitution of the form $({x_k\gets e_k})^{k\in K}$ 
	where $x_k$ are\ \ERIC{the state variables of $s'$}\footnote{all of them, completed by identity if needed}, and $e_k$ are expressions
        over the \ERIC{source or target state variables}\footnote{On pourrait du coup y mettre ici n'importe quelle variable du OA...}. Open transitions are identified
        modulo logical equivalence on their predicate. 
\end{definition}

It is important to understand the difference between the red dotted rule and a normal 
inference rule. They correspond to two different logical levels.
 The open transition is itself a logical implication, but using a simple logic (this logic includes the boolean expressions $\AlgB$, boolean operators, and term equality). \hl{We differentiate classical (black) inference rules that use  an expressive logic and are paper rules from open transition rules that use a simpler  logic,  will be embedded into automata transitions, and could typically be handled in a 
 mechanical way.}

@@

We take in this article a semantics and logical understanding of these
automata. Open automata are closed by a simple form of refinement that
allows us to refine the predicate, or substitute any free variable by
an expression. More formally, let $\Post$ be any substitution and
$\Pred$ any predicate; suppose $\vars(t)\cap \dom(\Post)=\emptyset$ and
$\vars(t')\cap \dom(\Post)=\emptyset$. Then we have the following
implication:

@@

\subsubsection*{Note on finite versus infinite open automata:} as
mentionned in page \pageref{def:open-automaton}, we have defined two flavors of
automata. More precisely, in \cite{hou:hal-02406098}, we formely define
\emph{semantic open automata} (infinite as in Definition \ref{def:open-automaton}),
and \emph{structural open automata} (finite) that can be generated as
the semantics of pNets (see Definition \ref{def:operationalSemantics}), and used in the implementation. Then we define
an alternative version of our bisimulation, called
structural-FH-Bisimulation, based on structural open automata, and
prove that the \emph{semantic} and \emph{structural} FH-Bisimulations coincide.
In the sequel, all mentions of finite automata, and algorithms for
bisimulations, implicitly refer to their \emph{structural} versions.

@@

\begin{theorem}[Decidability of FH-bisimulation]
Let $A_1$ and $A_2$ be finite open automata
and $\mathcal{R}$ a relation over their states $\mathcal{S}_1$ and
$\mathcal{S}_2$ constrained by a finite set of predicates. Assume that
the predicates inclusion is decidable over  
\hl{the action algebra} $\mathcal{A}_P$. Then it is decidable whether the relation 
$\mathcal{R}$ is a FH-bisimulation.

@@

  For any pNet \pNet, let $\Leaves(\pNet) = \mylangle S_i,{s_i}_0, \to_i\myrangle^{i \in L}$ be 
  the set of pLTS at its leaves,
  then $States(\pNet) = \{\triangleleft s_i^{i\in L}
  \triangleright| \forall i\in L. s_i \in S_i\}$.
A pLTS being its own single leave:
  $States(\mylangle S,s_0, \to\myrangle) = \{\triangleleft s \triangleright| s \in S, \}$. To be precise, the state of a pLTS is not reduced to a state identity, it also contains the value of its variables $vars(P)$. 
%This could be encoded as a mapping from variables of the pNet to expressions, of the form: $\sigma=\{x_i\to e_i\}^{i\in I},\,(x_i\in vars(S))$.
%\LUDO{j ai ajoute les variables mais c est pas tres beau}

@@

\begin{definition}[Semantics of open pNets]
	\label{def:operationalSemantics}
\hl{The semantics of a pNet} $\pNet$ is an open automaton $A\!= 
	<\!\!Holes(\pNet),States(\pNet),InitState(\pNet),vars(P),\mathcal{T}\!\!>$ where $\mathcal{T}$   is the smallest set of open transitions such that $\mathcal{T}=\{OT\,|\,\pNet \models OT \}$ and	$\pNet \models OT$	is defined by the following  rules:
%	\begin{itemize}
%		\item $J$ is the set of holes: $Holes(p)= J$. 
		%  \item $\set{L}^L = Leaves(p), \set{H}^J = Holes(p)$
%		\item ${\mathcal{S}} = States(p)$ and $s_0 = InitState(p)$
%		\item $\mathcal{T}$ is the smallest set of open transitions		satisfying the rules below:
%	\end{itemize}

@@

\paragraph{Example} \hl{Figure} \ref{SimpleProtCounter:SpecOA} shows the open automaton computed from the Specification pNet given in Figure \ref{SimpleProt:Spec}. 
For later references, we give to the transitions of this (strong)
Specification automaton as $SS_i$ while transitions of the
Implementation automaton will be labelled $SI_i$. In the figure we
have annotated each global state with the set of corresponding state variables.

>>>>

For this property to be valid even if the substitution does not operate on all variables,
we define the composition operation as follows: 
%\TODO{version d eric reprise et corrigee}
%$(x_k\gets e_k)^{k\in K}\shortotimes \Post =  (x_k\gets e_k\subst{\Post})^{k\in K}$
%\ERIC{J'ai un probleme ici... quand $Post'$ agit sur des vars de $t$ qui ne sont pas substitutees par $Post$, alors les 2 defs de $\shortotimes$ ci-dessus ne sont pas equivalentes:}\\
%$F(x,z)(\subst{x\gets z}\shortotimes\subst{z\gets 1})=F(x,z)(\subst{x\gets 1})=F(1,z)$\\
%\ERIC{Alors que}\\
%$(F(x,z)\subst{x\gets z})\subst{z\gets 1})=F(z,z)\subst{z\gets 1}=F(1,1)$
%
%\ERIC{Une definition plus correcte serait peut-etre:}\\

\[(x_{k}\gets e_{k})^{k\in K}\shortotimes (x'_{k'}\gets e'_{k'})^{k'\in K'} =  
(x_{k}\gets e_{k}\subst{(x'_{k'}\gets e'_{k'})^{k'\in K'}})^{k\in K} \cup (x'_{k'}\gets e'_{k'})^{k'\in K''}\]
where $K''=\{k'\in K'|x'_{k'}\not\in\{x_k\}^{k\in K}\}$
%
%\LUDO{En fait comme on supose qu il y a pas de variable globale et donc que les post doivent affecter toutes les variables ce contre-exemple n est pas applicable! apres on peut definir comme ca si on veut ceinture et bretelles mais ca alourdit un peu la definitions}
%
%\ERIC{Heu Non... 
%  En pratique, dans la composition de posts de la regle WT3, les posts assignent bien toutes les variables de leur OT (strong), mais dans la composition, on fini par avoir des trous... Voir page 21-22. On est tombes tous les deux separemment, Rabea et moi, sur ce genre de cas.}
% 
%\LUDO{euuh ca doit etre lie qu fait qu il manque des ec leftarrowec dans les selfloop de l'OA de spec. Plutot que changer la def et les preuves je propose de rajouter la condition $vars(s')=dom(Post)$ pour tous les OAs ce qui ne change rien a part pour les selfloop que l on s autorisait apparemment a ecrire en omettant $x:=x$ ... au pire l'identite est un post toujours correct et valide}

@@

\begin{definition}[pLTS]
\label{pLTS}
A pLTS is a tuple
$\pLTS\triangleq\mylangle S,s_0, V, \to\myrangle$ where:
\begin{itemize}
\item[$\bullet$]
$S$ is a set of states.
\item[$\bullet$]
$s_0 \in S$ is the initial state.
%\item[$\bullet$]
 %Variables in
%$\iv(\alpha)$ are assigned by the action, other variables can be assigned
%by the additional assignments.
\item[$\bullet$] $V$ is a set of global variables for the pLTS,
\item[$\bullet$] $\to \subseteq S \times L \times S$ is the transition relation and 
$L$ is the set of labels of the form
$\langle \alpha,~e_b,~(x_j\!:= {e}_j)^{j\in J}\rangle$,
where $\alpha \in\AlgA$ is a parameterised action, $e_b \in
\AlgB$ is a guard, and the variables $x_j$ 
are assigned the expressions $e_j\in \AlgE$.
If 
$s \xrightarrow{\langle \alpha,~e_b,~(x_j\!:= {e}_j)^{j\in
		J}\rangle} s'\in \to $ then 
% REMOVED BECAUSE USELESS: $\iv(\alpha)\!\subseteq\! \vars(s')$, 
		$\vars(\alpha)\backslash \iv(\alpha)\!\subseteq\! V$, 
		$\vars(e_b)\!\subseteq\! \vars(s)\cup\vars(\alpha)$, and
		$\forall j\!\in\! J .\,\left(\vars(e_j)\!\subseteq\! V\cup\vars(\alpha)\land 
		x_j\!\in V \right)$. %,  and $s= s'\Rightarrow J=\emptyset$.

%\LUDO{I suggest to add: $\{x_j|j\in J\} = \vars(s')$ or $\{x_j|j\in J\} \cup \iv(\alpha)= \vars(s')$} 
\end{itemize}
\end{definition}

@@

\item[$\bullet$] $\symb{SV}_k^{k\in K}$ is a set of
  synchronisation vectors. $\forall k\!\in\! K,
  \symb{SV}_k\!=\!\SV{\alpha_{l}^{l\in I_k \uplus J_k}}{\alpha'_k}{e_k}$ \hl{where}
  $\alpha'_k\in \AlgAS$, $I_k\subseteq I$, $J_k\subseteq J$,
  $\forall i\!\in\!
  I_k.\,\alpha_{i}\!\in\!\Sort(\pNet_i)$,  $\forall j\!\in\!
  J_k.\,\alpha_{j}\!\in\!\Sort_j$, and $\vars(\alpha'_k)\subseteq \bigcup_{l\in I_k\uplus 
  J_k}{\vars({\alpha_l})}$. The global action of a vector $\symb{SV}_k$ is
$\alpha'_k$. $e_k \in \AlgB$ is a guard associated to the vector such that
$\vars(e_k)\subseteq \bigcup_{l\in I_k\uplus J_k}{\vars({\alpha_l})}$.
\end{itemize}
Synchronisation vectors are identified modulo renaming of variables that appear in their 
action terms.
\end{definition}

@@

  \begin{itemize}
  \item The sort of a pNet is its signature, i.e. the set of actions in $\AlgAS$ it can
perform, where each action signature is an action 
label plus the arity of the action. In the definition of sorts, we do not need to 
distinguish
input variables and we remove the
\emph{input marker} (?) of variables.
\[
\begin{array}{l}
\Sortop(\mylangle S,s_0,V, \to\myrangle) = \{\alpha\subst{?x \gets x| 
x\in\symb{iv}(\alpha)}|s \xrightarrow{\langle \alpha,~e_b,~(x_j\!:= {e}_j)^{j\in
    J}\rangle} s'\in \to \} \\
\Sortop(\mylangle \set{\pNet}\!, %\pNet_i^{i\in I}, 
\set{\Sort},
\set{\symb{SV}}\myrangle)
=\{\alpha' |\, \SV{\set{\alpha}}{\alpha'}{e_b}\in\set{\symb{SV}}\}
\end{array}
\]

\item The set of variables of a pNet $P$, denoted $vars(P)$ is the set of variables of  all pLTSs that compose $P$.

\item
The set of holes $\Holes(\pNet)$ of a pNet is the indexes of the holes of the pNet 
itself plus the indexes of all the holes of its subnets.
It is defined inductively (we suppose those indexes 
disjoints):
  \[\begin{array}{l}
\Holes(\mylangle S,s_0,V, \to\myrangle) \!=\! \emptyset\\
\Holes(\mylangle \pNet_i^{i\in I}\!,\set{\Sort}, \set{\symb{SV}}\myrangle) 
=J\uplus{\displaystyle \bigcup_{i\in 
I}\Holes(\pNet_i)}\\
\forall i\in I.\, \Holes(\pNet_i)\cap J=\emptyset\\
\forall i_1,i_2\in I.\,i_1\neq i_2\Rightarrow  \Holes(\pNet_{i_1})\cap\Holes(\pNet_{i_2})=\emptyset
\end{array}\]
\item
The set of leaves of a pNet is the set of all pLTSs occurring in the structure, as an 
indexed family of the form $\Leaves(\pNet)= \mylangle \pNet_i \myrangle^{i \in L}$.
\[\begin{array}{l}
\Leaves(\mylangle S,s_0,V, \to\myrangle) \!=\!\emptyset\\%\{\mylangle S,s_0, \to\myrangle\}\\
\Leaves(\mylangle \pNet_i^{i\in I}\!,%Sort_j^{j\in J}
\set{\Sort}\!, \set{\symb{SV}}\myrangle) = {\displaystyle \biguplus_{i\in 
I}\Leaves(\pNet_i)\uplus\{i\mapsto \pNet_i|\pNet_i \text{ is a }\pLTS\}}
\end{array}\]
\end{itemize}

@@

\subsection{Open Automata}
 Open automata (OA) are not composition structures but they are made of transitions that are dependent of the actions of the holes, and they can reason on a set of variables (potentially with only symbolic values). 
%\TODO{adopt a uniform notation for open transitions, almost each instance has a 
%different 
%notation! I suggest p,l,pr,po using \{\} for p,l,po as they are sets}
\begin{definition}[Open transitions]\label{def:OT}
	\label{def:OpenTransitions}
	An \emph{open transition} (OT) over a
	set $J$ of holes with sorts $\Sort_j^{j\in J}$, a set V of variables, and a set of states $\mathcal{S}$ is 
	a structure of the form:	
	\begin{mathpar}
	\openrule
	{	\beta_j^{j\in J'}, \Pred, \Post}
	{s \OTarrow {\alpha}s'}
	\end{mathpar}
	Where $J'\subseteq J$, $s, s'\in\mathcal{S}$ and $\beta_j$
        is a transition of the hole $j$, with $\beta_j\in\Sort_j$. $\alpha$ is an action 
        label denoting the resulting action of this open transition.
%        We call \emph{source variables} of an OT the union of all variables
%        in the different terms
%        $\beta_j$ and $\alpha$, and the state variables of $s$.
        \Pred\ is a predicate over the  variables $V$ and all variables
        in the different terms
       $\beta_j$ and $\alpha$. \Post\ is a set of 
	assignments that are effective \emph{after the open transition}, they are
        represented as a substitution of the form $({x_k\gets e_k})^{k\in K}$ 
	where $\forall k.\, x_k\in V$, and $e_k$ are expressions
        over the variables $V$ and all variables
        in the different terms
       $\beta_j$ and $\alpha$. Open transitions are identified
        modulo logical equivalence on their predicate. 
\end{definition}

It is important to understand the difference between the red dotted rule and a normal 
inference rule. They correspond to two different logical levels.
On one side, classical (black) inference rules  use  an expressive logic and are paper rules.
 On the other side, open transition rules (with dotted lines) are logical implications, but using a simple logic (this logic includes the boolean expressions $\AlgB$, boolean operators, and term equality).
   open transitions could typically be handled in a 
 mechanised way.

@@

We take in this article a semantics and logical understanding of these
automata. Open automata are closed by a simple form of refinement that
allows us to refine the predicate, or substitute any free variable by
an expression. More formally, let $\Post$ be any substitution and
$\Pred$ any predicate; suppose $V\cap \dom(\Post)=\emptyset$. Then we have the following
implication:

@@

\subsubsection*{Note on finite versus infinite open automata:} 
As mentioned in Definition \pageref{def:open-automaton}, we adopt here a semantic view on open automata. More precisely, in \cite{hou:hal-02406098}, we formely define
\emph{semantic open automata} (infinite as in Definition \ref{def:open-automaton}),
and \emph{structural open automata} (finite) that can be generated as
the semantics of pNets (see Definition \ref{def:operationalSemantics}), and used in the implementation. Then we define
an alternative version of our bisimulation, called
structural-FH-Bisimulation, based on structural open automata, and
prove that the \emph{semantic} and \emph{structural} FH-Bisimulations coincide.
In the sequel, all mentions of finite automata, and algorithms for
bisimulations, implicitly refer to their \emph{structural} versions.

@@

\begin{theorem}[Decidability of FH-bisimulation]
Let $A_1$ and $A_2$ be finite open automata
and $\mathcal{R}$ a relation over their states $\mathcal{S}_1$ and
$\mathcal{S}_2$ constrained by a finite set of predicates. Assume that
the predicates inclusion is decidable over  
the action algebra $\AlgA$. Then it is decidable whether the relation 
$\mathcal{R}$ is a FH-bisimulation.

@@

  For any pNet \pNet, let $\Leaves(\pNet) = \mylangle S_i,{s_i}_0,V, \to_i\myrangle^{i \in L}$ be 
  the set of pLTS at its leaves,
  then $States(\pNet) = \{\triangleleft s_i^{i\in L}
  \triangleright| \forall i\in L. s_i \in S_i\}$.
A pLTS being its own single leave:
  $States(\mylangle S,s_0,V, \to\myrangle) = \{\triangleleft s \triangleright| s \in S, \}$. To be precise, the state of a pLTS is not reduced to a state identity, it also contains the value of its variables $vars(P)$. 
%This could be encoded as a mapping from variables of the pNet to expressions, of the form: $\sigma=\{x_i\to e_i\}^{i\in I},\,(x_i\in vars(S))$.
%\LUDO{j ai ajoute les variables mais c est pas tres beau}

@@

\begin{definition}[Semantics of open pNets]
	\label{def:operationalSemantics} The semantics of a pNet $\pNet$ is an open automaton $A\!= 
	<\!\!Holes(\pNet),States(\pNet),InitState(\pNet),vars(P),\mathcal{T}\!\!>$ where $\mathcal{T}$   is the smallest set of open transitions such that $\mathcal{T}=\{OT\,|\,\pNet \models OT \}$ and	$\pNet \models OT$	is defined by the following  rules:
%	\begin{itemize}
%		\item $J$ is the set of holes: $Holes(p)= J$. 
		%  \item $\set{L}^L = Leaves(p), \set{H}^J = Holes(p)$
%		\item ${\mathcal{S}} = States(p)$ and $s_0 = InitState(p)$
%		\item $\mathcal{T}$ is the smallest set of open transitions		satisfying the rules below:
%	\end{itemize}

@@

\paragraph{Example} \hl{paragraph} Figure~\ref{SimpleProtCounter:SpecOA} shows the open automaton computed from the Specification pNet given in Figure \ref{SimpleProt:Spec}. 
For later references, we give to the transitions of this (strong)
Specification automaton as $SS_i$ while transitions of the
Implementation automaton will be labelled $SI_i$. In the figure we
have annotated each global state with the set of corresponding state variables.
